---
title: "In-class Exercise 3: Working with Open Government Data"
author: "Dr. Kam Tin Seong<br/>Assoc. Professor of Information Systems(Practice)"
institute: "School of Computing and Information Systems,<br/>Singapore Management University"
date: "last-modified"
format: 
  revealjs:
    html: default
    width: 1600
    height: 900
    show-notes: false
    slide-number: true
    show-slide-number: all
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
editor: visual
---

## Learning Outcome

By the end of this hands-on exercise, you will be able to:

-   Preparing **ACRA (Accounting and Corporate Regulatory Authority) Information on Corporate Entities** datasets downloaded from data.gov.sg portal for geocoding,
-   Geocoding the tidydata by using SLA OneMap API,
-   Converting the geocoded transaction data into sf point feature data.frame, and
-   Wrangling the sf point features to avoid overlapping point features.

## Loading the R package

:::: panel-tabset
### The task

Write a code chunk to install and load tidyverse, sf, tmap and httr into R environment.

::: callout-note

- [**httr**](https://httr.r-lib.org/) is an R package specially designed to provide a wrapper for the [**curl**](https://cran.r-project.org/web/packages/curl/vignettes/intro.html) package, customised to the demands of modern web APIs. Key features: Functions for the most important http verbs: `GET()`, `HEAD()`, `PATCH()`, `PUT()`, `DELETE()` and `POST()`.
- [**dplyr**](https://dplyr.tidyverse.org/) is a grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges.  It is a must learned package for modern data scientists and data analysts.  Refer to this [chapter](https://r4ds.hadley.nz/data-transform) to learn more about dplyr.

- [**lubridate**](https://lubridate.tidyverse.org/) is an R package specially designed to handle date and date-time data type.  Refer to this [chapter](https://r4ds.hadley.nz/datetimes.html) to learn more about handle date and datetime with lubridate.
:::

### The code

::: {style="font-size: 1.50em"}
```{r}
pacman::p_load(tidyverse, sf, tmap, httr)
```
:::
::::

## Importing ACRA data

:::: panel-tabset
### The task

Write a code chunk to perform the followings:

- importing an ACRA data set into RStudio as tibble data frame, and 
- drop the unnecessary fields such as fields with excessive missing values or fields with only single values.

### The code

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
folder_path <- "data/aspatial/acra"
file_list <- list.files(path = folder_path, 
                        pattern = "^ACRA*.*\\.csv$", 
                        full.names = TRUE)

acra_data <- file_list %>%
  map_dfr(read_csv)
```
:::
::::

## Saving ACRA data

:::: panel-tabset
### The task

Write a code chunk to save the consolidated ACRA data.

### The code

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
write_rds(acra_data,
          "data/rds/acra_data.rds")
```
:::
::::

```{r}
#| echo: false
acra_data <- read_rds("data/rds/acra_data.rds")
```

## Tidying ACRA data

:::: panel-tabset
### The task
Write a code chunk to perform the followings:

- select the target businesses by using SSIC code,
- derive the year and month fields from *registration_incorporation_date* field,
- tidy postal code values to ensure avoid 5-digit postal codes by mistake, and
- select businesses registered in 2025.

### The code

::: {style="font-size: 1.25em"}
```{r}
#| eval: false
biz_56111 <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code == 56111) %>%
  rename(date = registration_incorporation_date) %>%
  mutate(date = as.Date(date),
         YEAR = year(date),
         MONTH_NUM = month(date),
         MONTH_ABBR = month(date, 
                            label = TRUE, 
                            abbr = TRUE)) %>% 
  mutate(
    postal_code = str_pad(postal_code, 
    width = 6, side = "left", pad = "0")) %>%
           filter(YEAR == 2025)    
```
:::
::::

## Geocoding

:::: panel-tabset
### The task

Write a function to perform the following tasks:

- Extracting unique postal code from the *postal_code* field of the tidied tibble data frame,
- performing geocode by submitting the postal codes to SLA OneMap [Reverse Geocode API](https://www.onemap.gov.sg/apidocs/reversegeocode/#svy21), and
- save the return results into two data frames namely *found* (for records that the postal codes are found) and *not_found* (for record that postal codes are not found). 

### The code
::: {style="font-size: 1.25em"}

```{r}
#| eval: false
postcodes <- unique(biz_56111$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

found <- data.frame()
not_found <- data.frame(postcode = character())

for (pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails = "Y",
    pageNum = "1"
  )
  
  res <- GET(url, query = query)
  json <- content(res)
  
  if (json$found != 0) {
    df <- as.data.frame(json$results, stringsAsFactors = FALSE)
    df$input_postcode <- pc
    found <- bind_rows(found, df)
  } else {
    not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
}
```
:::
::::

## Tidying the geocoded data

:::: panel-tabset

### The task

The geocoded data frame (i.e. found) contained many redundant fields.  We only need the first ten fields.  Write a code chunk to tidy the geocoded data frame.

### The code

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
found <- found %>%
  select(1:10)
```
:::
::::

## Appending the location information

:::: panel-tabset
### The taks

Write a code chunk to append the location information (i.e. x- and y-coordinates) to *biz_56111* sf data frame.

### The code

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
biz_56111 = biz_56111 %>%
  left_join(found, 
            by = c('postal_code' = 'POSTAL'))
```
:::
::::

## Saving the data

:::: panel-tabset
### The task
Write a code chunk to save the tidied data into rds format for future use.

### The code

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
write_rds(biz_56111, "data/rds/biz_56111.rds")
```
:::
::::

```{r}
#| echo: false
biz_56111 <- read_rds("data/rds/biz_56111.rds")
```

## Converting into SF data frame

:::: panel-tabset
### The Task

Write a code chunk to convert biz_56111 tibble data frame into sf data frame.

### The code

::: {style="font-size: 1.50em"}

```{r}
biz_56111_sf <- st_as_sf(biz_56111, 
                         coords = c("X","Y"),
                         crs=3414) 

```
:::
::::

## Visualising the distribution

:::: panel-tabset
### The task

Write a code chunk to reveal the distribution of newly registered businesses by month.

```{r}
#| echo: false
ggplot(data = biz_56111,
       aes(x = MONTH_ABBR)) +
  geom_bar()
```

### The code

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
ggplot(data = biz_56111,
       aes(x = MONTH_ABBR)) +
  geom_bar()
```
:::
::::

## Visualising the business

:::: panel-tabset
### The Task

```{r}
#| echo: false
tmap_mode('view')
tm_shape(biz_56111_sf)+
  tm_dots()
```


### The code

::: {style="font-size: 1.50em"}

```{r}
#| eval: false
tmap_mode('view')
tm_shape(biz_56111_sf)+
  tm_dots()
```
:::
::::


