---
title: "In-class Exercise 4: Geographically Weighted Summary Statistics - gwModel methods"
author: "Dr. Kam Tin Seong<br/>Assoc. Professor of Information Systems(Practice)"
institute: "School of Computing and Information Systems,<br/>Singapore Management University"
date: "`r Sys.Date()`"
format: 
  revealjs:
    pdf: default
    width: 1600
    height: 900
    show-notes: false
    slide-number: true
    show-slide-number: all
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
date-format: "DD MMM YYYY"
editor: visual 
---

## Loading the package

In this in-class exercise, sf, spdep, tmap, tidyverse, knitr and GWmodel will be used.

:::: panel-tabset
### DIY

Using the step you leanred from previous hands-in, install and load the necessary R packages in R environment.

### The code

::: {style="font-size: 1.50em"}
```{r}
pacman::p_load(sf, ggstatsplot, tmap, tidyverse, knitr, GWmodel)
```
:::
::::

## Preparing the Data

For this in-class exercise, Hunan shapefile and Hunan_2012 data file will be used.

:::::: panel-tabset
### DIY

Using the steps you learned from previous hands-on, complete the following tasks:

-   import Hunan shapefile and parse it into a sf polygon feature object.
-   import Hunan_2012.csv file parse it into a tibble data.frame.
-   join Hunan and Hunan_2012 data.frames.

### Importing Hunan shapefile

::: {style="font-size: 1.50em"}
```{r}
hunan_sf <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```
:::

### Importing Hunan_2012 table

::: {style="font-size: 1.50em"}
```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```
:::

### Joining Hunan and Hunan_2012

::: {style="font-size: 1.50em"}
```{r}
hunan_sf <- left_join(hunan_sf, hunan2012) %>%
  select(1:3, 7, 15, 16, 31, 32)
```
:::
::::::

## Mapping GDPPC

:::: panel-tabset
### DIY

Using the steps you learned from Hands-on Exercise 5, prepare a choropleth map showing the geographic distribution of GDPPC of Hunan Province.

```{r}
#| fig-width: 16
#| fig-height: 6
#| echo: false
basemap <- tm_shape(hunan_sf) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan_sf, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

### The code

::: {style="font-size: 1.50em"}
```{r}
#| fig-width: 12
#| fig-height: 8
#| eval: false
basemap <- tm_shape(hunan_sf) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan_sf, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```
:::
::::

## Converting to SpatialPolygonDataFrame

::: callout-note
GWmodel presently is built around the older sp and not sf formats for handling spatial data in R.
:::

::: {style="font-size: 1.50em"}
```{r}
hunan_sp <- hunan_sf %>%
  as_Spatial()
```
:::

## Geographically Weighted Summary Statistics (GSWW) with adaptive bandwidth

### Determine adaptive bandwidth

::::: panel-tabset
### Cross-validation

::: {style="font-size: 1.20em"}
```{r}
bw_CV <- bw.gwr(GDPPC ~ 1, 
             data = hunan_sp,
             approach = "CV",
             adaptive = TRUE, 
             kernel = "bisquare", 
             longlat = T)
bw_CV
```
:::

### AIC

::: {style="font-size: 1.20em"}
```{r}
bw_AIC <- bw.gwr(GDPPC ~ 1, 
             data = hunan_sp,
             approach ="AIC",
             adaptive = TRUE, 
             kernel = "bisquare", 
             longlat = T)
bw_AIC
```
:::
:::::

---

### Computing geographically wieghted summary statistics

The code chunk below computes geographically weighted summary statistics by using `gwss()` of GWmodel package.  The analysis is performed using adaptive distance weighted.  The distance is derived using AIC method and the kernel method used id bisquare.

::: {style="font-size: 1.50em"}
```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_AIC,
               kernel = "bisquare",
               quantile = TRUE,
               adaptive = TRUE,
               longlat = T)
```
:::

::: callout-note
gwss support five kernel methods, namely gaussian, exponential, bisquare, tricube, and boxcar.  Repeat the code chunk above with different kernel methods and compare their results visually.
:::

---

### Preparing the output data

::: callout-important
`gwss()` with **quantile** argument computes eight local summary statistics, namely: local means, local standard deviations, local variance, local skew, local coefficients of variation, local Median, local medians, local interquartile ranges, and local quantile imbalances.  They are output in **SDF** data frame.
:::

Code chunk below is used to extract **SDF** data table from **gwss** object output from `gwss()` for subsequent geovilisation. It will be converted into data.frame by using `as.data.frame()`.

::: {style="font-size: 1.50em"}
```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```
:::

Next, `cbind()` is used to append the newly derived data.frame onto *hunan_sf* sf data.frame.

::: {style="font-size: 1.50em"}
```{r}
hunan_gstat <- cbind(hunan_sf, gwstat_df)
```
:::

---

### Visualising geographically weighted summary statistics

:::::: panel-tabset
### The Geographically Weighted Mean

:::: columns
::: {.column width="50%"}
```{r}
#| fig-height: 9
#| echo: false
tm_shape(hunan_gstat) +
  tm_polygons(fill = "GDPPC_LM",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "local Mean")) +
  tm_borders(fill_alpha = 0.5) +
  tm_layout(main.title = "Distribution of geographically wieghted mean",
            main.title.position = "center",
            main.title.size = 2.0,
            frame = TRUE)
```
:::
::::

### The code

::: {style="font-size: 1.30em"}
```{r}
#| eval: false
tm_shape(hunan_gstat) +
  tm_polygons(fill = "GDPPC_LM",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "local Mean")) +
  tm_borders(fill_alpha = 0.5) +
  tm_layout(main.title = "Distribution of geographically wieghted mean",
            main.title.position = "center",
            main.title.size = 2.0,
            frame = TRUE)
```
:::
::::::


## GWSS with Fixed Bandwidth
### Determine fixed bandwidth

::::: panel-tabset
### Cross-validation

::: {style="font-size: 1.20em"}
```{r}
bw_CV <- bw.gwr(GDPPC ~ 1, 
             data = hunan_sp,
             approach = "CV",
             adaptive = FALSE, 
             kernel = "bisquare", 
             longlat = T)
bw_CV
```
:::

### AIC

::: {style="font-size: 1.20em"}
```{r}
bw_AIC <- bw.gwr(GDPPC ~ 1, 
             data = hunan_sp,
             approach ="AIC",
             adaptive = FALSE, 
             kernel = "bisquare", 
             longlat = T)
bw_AIC
```
:::
:::::

---

### Computing adaptive bandwidth GWSS

::: {style="font-size: 1.50em"}
```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_AIC,
               kernel = "bisquare",
               adaptive = FALSE,
               quantile = FALSE,
               longlat = T)
```
:::

---

### Preparing the output data

Code chunk below is used to extract **SDF** data table from **gwss** object output from `gwss()`. It will be converted into data.frame by using `as.data.frame()`.

::: {style="font-size: 1.50em"}
```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```
:::

Next, `cbind()` is used to append the newly derived data.frame onto *hunan_sf* sf data.frame.

::: {style="font-size: 1.50em"}
```{r}
hunan_gstat <- cbind(hunan_sf, gwstat_df)
```
:::

---

### Visualising geographically weighted summary statistics

:::::: panel-tabset
### The Geographically Weighted Mean

:::: columns
::: {.column width="50%"}
```{r}
#| fig-height: 9
#| echo: false
tm_shape(hunan_gstat) +
  tm_polygons(fill = "GDPPC_LM",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "local Mean")) +
  tm_borders(fill_alpha = 0.5) +
  tm_layout(main.title = "Distribution of geographically wieghted mean",
            main.title.position = "center",
            main.title.size = 2.0,
            frame = TRUE)
```
:::
::::

### The code

::: {style="font-size: 1.30em"}
```{r}
#| eval: false
tm_shape(hunan_gstat) +
  tm_polygons(fill = "GDPPC_LM",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "local Mean")) +
  tm_borders(fill_alpha = 0.5) +
  tm_layout(main.title = "Distribution of geographically wieghted mean",
            main.title.position = "center",
            main.title.size = 2.0,
            frame = TRUE)
```
:::
::::::

## Geographically Weighted Correlation

Business question: Is there any relationship between GDP per capita and Gross Industry Output?

:::: panel-tabset
### Conventional statistical solution

```{r}
#| echo: false
ggscatterstats(
  data = hunan2012, 
  x = GIO, 
  y = GDPPC,
  xlab = "Gross Industry Output", ## label for the x-axis
  ylab = "GDP per capita", 
  label.var = County, 
  label.expression = Agri > 10000 & GDPPC > 50000, 
  point.label.args = list(alpha = 0.7, size = 4, color = "grey50"),
  xfill = "#CC79A7", 
  yfill = "#009E73", 
  title = "Relationship between GDP PC and Gross Agriculture Output")
```

### The code

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
ggscatterstats(
  data = hunan2012, 
  x = Agri, 
  y = GDPPC,
  xlab = "Gross Agriculture Output", ## label for the x-axis
  ylab = "GDP per capita", 
  label.var = County, 
  label.expression = Agri > 10000 & GDPPC > 50000, 
  point.label.args = list(alpha = 0.7, size = 4, color = "grey50"),
  xfill = "#CC79A7", 
  yfill = "#009E73", 
  title = "Relationship between GDP PC and Gross Agriculture Output")
```
:::
::::

---

### Geospatial analytics solution

::::::: panel-tabset
### Determine the bandwidth

::: {style="font-size: 1.50em"}
```{r}
bw <- bw.gwr(GDPPC ~ GIO, 
             data = hunan_sp, 
             approach = "AICc", 
             adaptive = TRUE)
```
:::

### Computing gwCorrelation

::: {style="font-size: 1.50em"}
```{r}
gwstats <- gwss(hunan_sp, 
                vars = c("GDPPC", "GIO"), 
                bw = bw,
                kernel = "bisquare",
                adaptive = TRUE, 
                longlat = T)
```
:::

### Extracting the result

Code chunk below is used to extract **SDF** data table from **gwss** object output from `gwss()`. It will be converted into data.frame by using `as.data.frame()`.

::: {style="font-size: 1.50em"}
```{r}
gwstat_df <- as.data.frame(gwstats$SDF) %>%
  select(c(12,13)) %>%
  rename(gwCorr = Corr_GDPPC.GIO,
         gwSpearman = Spearman_rho_GDPPC.GIO)
```
:::

Next, `cbind()` is used to append the newly derived data.frame onto *hunan_sf* sf data.frame.

::: {style="font-size: 1.50em"}
```{r}
hunan_Corr <- cbind(hunan_sf, gwstat_df)
```
:::
:::::::

---

### Visualising Local Correlation

:::: panel-tabset
### Local Correlation Coefficient

```{r}
#| fig-height: 7
#| echo: false
tm_shape(hunan_Corr) + 
tm_polygons(fill = "gwCorr",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
                fill.legend = tm_legend(
                title = "local Correlation Coefficient")) +
  tm_borders(fill_alpha = 0.5) + 
  tm_title("Local Correlation Coefficient", 
  size = 2.0) +
  tm_layout(frame = TRUE)
```

### Local Spearman Coefficient

```{r}
#| fig-height: 7
#| echo: false
tm_shape(hunan_Corr) +
  tm_polygons(fill = "gwSpearman",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "local Spearman Rho")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Local Spearman Rho", 
  size = 2.0) +
  tm_layout(frame = TRUE)
```

### The code

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
tm_shape(hunan_Corr) +
  tm_polygons(fill = "gwSpearman",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "local Spearman Rho")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Local Spearman Rho", 
  size = 2.0) +
  tm_layout(frame = TRUE)
```
:::
::::

## References

Brunsdon, C. et. al. (2002) "Geographically weighted summary statistics - a framework for localised exploratory data analysis", *Computer, Environment and Urban Systems*, Vol 26, pp. 501-525. Available as e-journal, SMU library.

Harris, P. & Brunsdon, C. (2010) "Exploring spatial variation and spatial relationships in freshwater acidification critical load data set for Great Britain using geographically weighted summary statistics", *Computers & Geosciences*, Vol. 36, pp. 54-70. Available as e-journal, SMU library.

```{r}
#| eval: false
#| echo: false
renderthis::to_pdf(from ="In-class_Ex04.qmd")
```
